---
title: "Tidyverse Examples"
author: "Psych 251 Staff"
date: "10/2/2019"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

# Manipulating data with dplyr

Let's use `mtcars`, a built in dataset of cars and their miles/gallon (mpg), number of cylinders (cyl), displacement (disp), gross horsepower (hp), etc. 

```{r}
mtcars
```

**Exercise**: First, summarise the average miles/gallon (mpg) across the entire dataset. 

```{r}
mtcars %>%
  summarise(mean = mean(mpg))
```

**Exercise**: A car can either have 4, 6, or 8 cylinders (cyl). Summarise the average mpg, broken down by the number of cylinders. Hint: You may want to "group" by cyl in order to do this. 

```{r}
mtcars %>%
  group_by(cyl) %>%
  summarise(mean = mean(mpg))
```

**Exercise**: In addition to the means, add standard deviations to this summary (still grouped by cyl).

```{r}
mtcars %>%
  group_by(cyl) %>%
  summarise(mean = mean(mpg), 
            sd = sd(mpg))

```

**BONUS**: Let's visualize! Use ggplot (included in the tidyverse package) to make a scatter plot of mpg by horsepower. If you are feeling extra fancy, you can add a smoothing line. (Hint: Google "geom_smooth() scatterplot".)

```{r}
ggplot(mtcars, 
       aes(x = hp, y = mpg)) + 
  geom_point() + 
  geom_smooth() 
```



# Reshaping with tidyr

## From long to wide and back again

Use data from informative communication experiment. 

http://langcog.stanford.edu/papers/FG-cogpsych2014.pdf

```{r}
info <- read_csv("https://raw.githubusercontent.com/mcfrank/info_kid/master/data/info_e2_data.csv")
```

`info` is in tidy format. Make this into wide data. 

```{r}
info_wide <- info %>%
  spread(item, correct) 
```

Now make it back into tidy data. 

```{r}
info_long <- info_wide %>%
  gather(item, correct, bear:rocket)
```

## From wide to long without seeing the tidy version

# From wide to long without seeing the tidy version

These are pre-post data on children's arithemtic scores from an RCT in which they were assigned either to CNTL (control) or MA (mental abacus math intervention). They were tested twice, once in 2015 and once in 2016. 

https://jnc.psychopen.eu/article/view/106

```{r}
majic <- read_csv("data/majic.csv")
```

Make these tidy. 

```{r}
majic %>%
  gather(year, score, `2015`, `2016`)
```

